Reconstructing the neutron-star equation of state with gravitational-wave detectors from a realistic population of inspiralling binary neutron stars

### 2014-08-13 paper.tex
Added more to Section II

The strain $h$ observed in a gravitational-wave detector from a wave emitted by a BNS system is related to the two polarizations of the wave and to the detector's antenna beam pattern response:
\begin{equation}
h=F_+ h_+ + F_\times h_\times.
\end{equation}
For a single interferometric detector, the antenna patterns take the form
\begin{align}
\begin{split}
F_+(\theta,\phi,\psi)=&-\frac{1}{2}(1+\cos^2\theta)\cos(2\phi)\cos(2\psi)\\
&-\cos(\theta)\sin(2\phi)\sin(2\psi),
\end{split}\\
\begin{split}
F_\times(\theta,\phi,\psi)=&+\frac{1}{2}(1+\cos^2\theta)\cos(2\phi)\sin(2\psi)\\
&-\cos(\theta)\sin(2\phi)\cos(2\psi),
\end{split}
\end{align}
in a coordinate system where the two arms define the x-axis and the y-axis, and the z-axis is normal to the x-y plane.  Here, $(\theta, \phi)$ are the sky position in spherical coordinates, and $\psi$ is the polarization angle.  Assuming the binary components are point particles in a circular orbit, the two polarizations of the emitted gravitational wave seen by an observer at a distance $r$ away are
\begin{eqnarray}
\label{hoftp}
h_+(t)&=&-\frac{4\mu}{r}\left(\frac{1+\cos^2\iota}{2}\right)x(t)\cos\left[2\Phi(t)\right],\\
\label{hoftc}
h_\times(t)&=&-\frac{4\mu}{r}\cos\iota\mbox{ }x(t)\sin\left[2\Phi(t)\right],
\end{eqnarray}
in the time-domain, where $\mu=m_1 m_2/M$ is the reduced mass, $M=m_1+m_2$ is the total mass, $\iota$ is the inclination angle, $\Phi(t)$ is the orbital phase, and $x=(\pi M f)^{2/3}$ where $f$ is the gravitational-wave frequency.  The stationary-phase approximation can be used to find the two polarizations in the frequency-domain:
\begin{eqnarray}
\label{hoffp}
\tilde{h}_+(f)&=&\sqrt{\frac{5}{24}}\frac{{\cal M}^{5/6}f^{-7/6}}{\pi^{2/3}r}\left(\frac{1+\cos^2\iota}{2}\right)e^{i\Psi(f)},\\
\label{hoffc}
\tilde{h}_\times(f)&=&i\sqrt{\frac{5}{24}}\frac{{\cal M}^{5/6}f^{-7/6}}{\pi^{2/3}r}\cos\iota\mbox{ }e^{i\Psi(f)},
\end{eqnarray}
where the chirp mass ${\cal M}=\eta^{3/5}M$, the symmetric mass ratio $\eta=m_1 m_2/M^2$, $\Psi(f)=2\pi f t(v) - 2\Phi(v) - \pi/4$, and $\tilde{h}_+(f)$ and $\tilde{h}_\times(f)$ are the Fourier transforms of $h_+(t)$ and $h_\times(t)$ respectively.

@@

As was done in most previous works~\cite{WadeCreightonOchsner2014, DelPozzoLiAgathos2013, Favata2014, YagiYunes2014}, we keep the PN point-particle corrections to the energy and luminosity up to 3.5PN order as well as the leading-order and next-to-leading-order tidal contributions, which enter at the same order as 5PN and 6PN point-particle terms respectively.  The energy and luminosity take the form:
\begin{align}
E &= - \frac{1}{2} M \eta x \left[1 + e_\text{PP-PN}(\eta) + e_\text{Tidal}(\eta, \Lambda_1, \Lambda_2)\right], \\
\mathcal{L} &= \frac{32}{5}\eta^2 x^5 \left[ 1 + l_\text{PP-PN}(\eta) + l_\text{Tidal}(\eta, \Lambda_1, \Lambda_2)\right],
\end{align}
where the point-particle PN terms $e_\text{PP-PN}$ and $l_\text{PP-PN}$ are given in Ref.~\cite{Blanchet2014Review}, and the tidal terms are
\begin{widetext}
\begin{align}
\begin{split}
e_\text{Tidal} &= -\frac{9\eta}{2} \left[ (1-3\eta)(\Lambda_1+\Lambda_2) + \sqrt{1-4\eta}(1-\eta)(\Lambda_1-\Lambda_2) \right]x^5 \\
& - 22\eta \left[ \left(1-4\eta+\frac{19\eta^2}{8}\right)(\Lambda_1+\Lambda_2) + \sqrt{1-4\eta}\left(1-2\eta+\frac{3\eta^2}{8}\right)(\Lambda_1-\Lambda_2) \right]x^6,
\end{split} \\
\begin{split}
l_\text{Tidal} &= 3 \left[ (1-2\eta-4\eta^2)(\Lambda_1+\Lambda_2) + \sqrt{1-4\eta}(1-2\eta^2)(\Lambda_1-\Lambda_2) \right]x^5 \\
& - \frac{22}{7} \left[ \left(1-\frac{15\eta}{176}+\frac{147\eta^2}{88}-\frac{1547\eta^3}{44}\right)(\Lambda_1+\Lambda_2) + \sqrt{1-4\eta}\left(1+\frac{337\eta}{176}+\frac{15\eta^2}{2}-\frac{1085\eta^3}{88}\right)(\Lambda_1-\Lambda_2) \right]x^6,
\end{split}
\end{align}
\end{widetext}
where $\Lambda_1=\lambda_1/m_1$ and $\Lambda_2=\lambda_2/m_2$.  We have also asserted that $m_1 \ge m_2$ in order to express each correction in terms of dimensionless quantities.

@@

The majority of this work uses the TaylorF2 waveform approximant because its fully analytic frequency-domain form makes it the fastest PN waveform to generate.  For reference, the tidal corrections to the TaylorF2 phase are
\begin{widetext}
\begin{equation}
\delta\psi_{\rm tidal}=\frac{3}{128\eta x^{5/2}}\left[ \left(-\frac{39}{2}\tilde{\Lambda}\right)x^5+ \left(-\frac{3115}{64}\tilde{\Lambda}+\frac{6595}{364}\sqrt{1-4\eta}\mbox{ }\delta\tilde{\Lambda}\right)x^6 \right],
\end{equation}
where
\begin{eqnarray}
\label{LT}
\tilde{\Lambda}&=&\frac{8}{13}\left[\left(1+7\eta-31\eta^2\right)\left(\Lambda_1+\Lambda_2\right)+\sqrt{1-4\eta}\left(1+9\eta-11\eta^2\right)\left(\Lambda_1-\Lambda_2\right)\right]\\
\label{dLT}
\delta\tilde{\Lambda}&=&\frac{1}{2}\left[\sqrt{1-4\eta}\left(1-\frac{13272}{1319}\eta+\frac{8944}{1319}\eta^2\right)\left(\Lambda_1+\Lambda_2\right) + \left(1-\frac{15910}{1319}\eta+\frac{32850}{1319}\eta^2+\frac{3380}{1319}\eta^3\right)\left(\Lambda_1-\Lambda_2\right)\right].
\end{eqnarray}
\end{widetext}
For more details regarding this choice of tidal deformability parameters, see \cite{WadeCreightonOchsner2014} and \cite{Favata2014}.

>>>>

The strain $h$ observed in a gravitational-wave detector from a wave emitted by a BNS system is related to the two polarizations of the wave and to the detector's antenna beam pattern response:
\begin{equation}
h=F_+ h_+ + F_\times h_\times.
\end{equation}
For a single interferometric detector, the antenna patterns take the form
\begin{align}
\begin{split}
F_+(\theta,\phi,\psi)=&-\frac{1}{2}(1+\cos^2\theta)\cos(2\phi)\cos(2\psi)\\
&-\cos(\theta)\sin(2\phi)\sin(2\psi),
\end{split}\\
\begin{split}
F_\times(\theta,\phi,\psi)=&+\frac{1}{2}(1+\cos^2\theta)\cos(2\phi)\sin(2\psi)\\
&-\cos(\theta)\sin(2\phi)\cos(2\psi),
\end{split}
\end{align}
in a coordinate system where the two arms define the x-axis and the y-axis, and the z-axis is normal to the x-y plane.  Here, $(\theta, \phi)$ are the sky position in spherical coordinates, and $\psi$ is the polarization angle.  Assuming the binary components are point particles in a circular orbit, the two polarizations of the emitted gravitational wave seen by an observer at a distance $D$ away are
\begin{eqnarray}
\label{hoftp}
h_+(t)&=&-\frac{4\mu}{D}\left(\frac{1+\cos^2\iota}{2}\right)x(t)\cos\left[2\Phi(t)\right],\\
\label{hoftc}
h_\times(t)&=&-\frac{4\mu}{D}\cos\iota\mbox{ }x(t)\sin\left[2\Phi(t)\right],
\end{eqnarray}
in the time-domain, where $\mu=m_1 m_2/M$ is the reduced mass, $M=m_1+m_2$ is the total mass, $\iota$ is the inclination angle, $\Phi(t)$ is the orbital phase, and $x=(\pi M f)^{2/3}$ where $f$ is the gravitational-wave frequency.  The stationary-phase approximation can be used to find the two polarizations in the frequency-domain:
\begin{eqnarray}
\label{hoffp}
\tilde{h}_+(f)&=&\sqrt{\frac{5}{24}}\frac{{\cal M}^{5/6}f^{-7/6}}{\pi^{2/3}D}\left(\frac{1+\cos^2\iota}{2}\right)e^{i\Psi(f)},\\
\label{hoffc}
\tilde{h}_\times(f)&=&i\sqrt{\frac{5}{24}}\frac{{\cal M}^{5/6}f^{-7/6}}{\pi^{2/3}D}\cos\iota\mbox{ }e^{i\Psi(f)},
\end{eqnarray}
where the chirp mass ${\cal M}=\eta^{3/5}M$, the symmetric mass ratio $\eta=m_1 m_2/M^2$, $\Psi(f)=2\pi f t(v) - 2\Phi(v) - \pi/4$, and $\tilde{h}_+(f)$ and $\tilde{h}_\times(f)$ are the Fourier transforms of $h_+(t)$ and $h_\times(t)$ respectively.

@@

As was done in most previous works~\cite{WadeCreightonOchsner2014, DelPozzoLiAgathos2013, Favata2014, YagiYunes2014}, we keep the PN point-particle corrections to the energy and luminosity up to 3.5PN order as well as the leading-order and next-to-leading-order tidal contributions, which enter at the same order as 5PN and 6PN point-particle terms respectively.  The energy and luminosity take the form:
\begin{align}
E &= - \frac{1}{2} M \eta x \left[1 + e_\text{PP-PN}(\eta) + e_\text{Tidal}(\eta, \Lambda_1, \Lambda_2)\right], \\
\mathcal{L} &= \frac{32}{5}\eta^2 x^5 \left[ 1 + l_\text{PP-PN}(\eta) + l_\text{Tidal}(\eta, \Lambda_1, \Lambda_2)\right],
\end{align}
where the point-particle PN terms $e_\text{PP-PN}$ and $l_\text{PP-PN}$ are given in Ref.~\cite{Blanchet2014Review}, and the tidal terms are
\begin{widetext}
\begin{eqnarray}
e_\text{Tidal} &=& -\frac{9\eta}{2} \left[ (1-3\eta)(\Lambda_1+\Lambda_2) + \sqrt{1-4\eta}(1-\eta)(\Lambda_1-\Lambda_2) \right]x^5 \\
&& - 22\eta \left[ \left(1-4\eta+\frac{19\eta^2}{8}\right)(\Lambda_1+\Lambda_2) + \sqrt{1-4\eta}\left(1-2\eta+\frac{3\eta^2}{8}\right)(\Lambda_1-\Lambda_2) \right]x^6,\\
l_\text{Tidal}&=&3 \left[ (1-2\eta-4\eta^2)(\Lambda_1+\Lambda_2) + \sqrt{1-4\eta}(1-2\eta^2)(\Lambda_1-\Lambda_2) \right]x^5 \\
&& - \frac{22}{7} \left[ \left(1-\frac{15\eta}{176}+\frac{147\eta^2}{88}-\frac{1547\eta^3}{44}\right)(\Lambda_1+\Lambda_2) \right . \\
&& \left . + \sqrt{1-4\eta}\left(1+\frac{337\eta}{176}+\frac{15\eta^2}{2}-\frac{1085\eta^3}{88}\right)(\Lambda_1-\Lambda_2) \right]x^6,
\end{eqnarray}
\end{widetext}
where $\Lambda_1=\lambda_1/m_1$ and $\Lambda_2=\lambda_2/m_2$.  We have also asserted that $m_1 \ge m_2$ in order to express each correction in terms of dimensionless quantities.

@@

The majority of this work uses the TaylorF2 waveform approximant because its fully analytic frequency-domain form makes it the fastest PN waveform to generate.  The complete TaylorF2 strain can be written more compactly in the following form:
\begin{equation}
\tilde{h}(f)=\sqrt{\frac{5}{24}}\frac{{\cal M}^{5/6}}{\pi^{2/3}}\frac{Q(\theta,\phi,\psi,\iota)}{D}f^{-7/6}e^{i\Psi(f)},
\end{equation}
where
\begin{equation}
Q(\theta,\phi,\psi,\iota) = F_+(\theta,\phi,\psi)\frac{1+\cos\iota}{2} + i F_\times(\theta,\phi,\psi)\cos\iota.
\end{equation}
For a single detector, the extrinsic parameters in the amplitude can be grouped together into a single effective distance parameter, which is $D_{\rm eff}=|Q(\theta,\phi,\psi,\iota)|/D$.  The analytic form for the phase $\Psi(f)$ with point-particle terms up to 3.5PN order and leading-order and next-to-leading-order tidal contributions is
\begin{equation}
\Psi(f)=\frac{3}{128\eta x^{5/2}}\left[1+\psi_\text{PP-PN}(\eta)+\psi_\text{Tidal}(\eta,\Lambda_1,\Lambda_2)\right],
\end{equation}
where $\psi_\text{PP-PN}(\eta)$ is given in \cite{BuonannoIyerOchsner2009}, and
\begin{widetext}
\begin{equation}
\label{psi_tidal}
\psi_\text{Tidal}(\eta,\Lambda_1,\Lambda_2)=\left[-\frac{39}{2}\tilde{\Lambda}(\eta,\Lambda_1,\Lambda_2)\right]x^5+ \left[-\frac{3115}{64}\tilde{\Lambda}(\eta,\Lambda_1,\Lambda_2)+\frac{6595}{364}\sqrt{1-4\eta}\mbox{ }\delta\tilde{\Lambda}(\eta,\Lambda_1,\Lambda_2)\right]x^6,
\end{equation}
where
\begin{eqnarray}
\label{LT}
\tilde{\Lambda}(\eta,\Lambda_1,\Lambda_2)&=&\frac{8}{13}\left[\left(1+7\eta-31\eta^2\right)\left(\Lambda_1+\Lambda_2\right)+\sqrt{1-4\eta}\left(1+9\eta-11\eta^2\right)\left(\Lambda_1-\Lambda_2\right)\right],\\
\label{dLT}
\delta\tilde{\Lambda}(\eta,\Lambda_1,\Lambda_2)&=&\frac{1}{2}\left[\sqrt{1-4\eta}\left(1-\frac{13272}{1319}\eta+\frac{8944}{1319}\eta^2\right)\left(\Lambda_1+\Lambda_2\right) \right .\\
&& + \left . \left(1-\frac{15910}{1319}\eta+\frac{32850}{1319}\eta^2+\frac{3380}{1319}\eta^3\right)\left(\Lambda_1-\Lambda_2\right)\right].
\end{eqnarray}
\end{widetext}
We express the tidal contributions to the TaylorF2 phase in terms of $(\tilde\Lambda,\delta\tilde\Lambda)$ because $\tilde\Lambda$ is comparably more measurable than $\Lambda_1$ and $\Lambda_2$ individually.  For more details regarding this parameterization, see \cite{WadeCreightonOchsner2014,Favata2014}.
